---
title: "w3_Iterations"
author: "Amy Yi-Ting Lu"
date: "9/17/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# For loops

```{r}

x <- c(1, 3, 1, 4, 5 ,2, 10)

sumval <- 0 ## initial sum of a vector ## initiallize with 0 # you can also use length(), but seq_along is better

for (i in seq_along(x)){
  sumval <- sumval +x[i]
}

sumval
sum(x)

```

Always initialize the output before the for-loop

In the above for-loop, 

the **output**: 'sumval'
the **sequence**: seq_along(x) - the possible values of "i"
the **body**: 'sumval <- sumval + x[i]' -- what happend each interation as "i" changes.


Calculate the cumulative summations of a vector

```{r}

## 3, 4, 5, 6, 8, 2 --> accumulative vector will be 3 ,7 ,12 ,18, 26, 28

x
cumsum(x)

## initialize the output

cumvec <- vector(mode = "numeric", length = length(x))
cumvec

for(index in seq_along(x)){
  if (index == 1) {
    cumvec[1] <- x[index]
  } else {
  cumvec[index] <- cumvec[index -1] + x[index]
}
}

cumvec


```

Exercise

```{r}

fnum <- c(1:100)
fnum[1] <- 0
fnum[2] <- 1
fnum[3] <- 1

fnum


for (i in seq_along(x)) {
  if(index == 1){
    fnum[1] <- x[index]
  } else if (index == 2){
    fnum[2] <- x[index -1]
  }else if (index == 3){
    fnum[3] <- x[index -2]
  }else{
    fnum[index] <- fnum[index -3] + x[index]
  }
  
}


```









