---
title: "w3_Vector"
author: "Amy Yi-Ting Lu"
date: "9/17/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Vector Basic

```{r}

x <- c(1, 2, 10)

typeof(x)

## double means number can be with decimal point while integer can only be 1, 2, or 4...etc

```

Need to load the tidyverse for "is_" function

```{r}

library(tidyverse)

is_double(x) ## return TRUE
is_double("hello") ## it is not double so it returns fault

```

Follow numbers by "L" to specify an integer

```{r}

x <- c(1L, 2L, 10L)

is_double(x)
typeof(x)

```


```{r}

two <- sqrt(2) ^2
two

typeof(two)
two == 2 ## this 2 is actually just a run up of 2 (very close to 2)

```

```{r}

y <- c("2", "hello", "two")

typeof(y)

is_character(y)

```

```{r}

fvec <- parse_factor(c("A", "B", "A", "B", "C"))
fvec


typeof(fvec)
attributes(fvec)


```

Create a name vector by having the name to the left of an equal sign and the value to the right

```{r}

name_vec <- c(horse = 7, man = 1, dog = 8)
name_vec
typeof(name_vec)

```

Extract these name using 'names()' function

```{r}

names(name_vec)


```

Also use 'names' function to change the names of an observation

```{r}

names(name_vec)[1] <- "cat"
name_vec

```

You can extract elements using names
```{r}

name_vec["cat"]

```

OTher ways to subset: specifying the indice of the lcoation of the observations:

```{r}

x <- c("I", "like", "dogs")
x[2:3]
x[c(2:3)]
x[c(1,3)]

```

You can specify subeets with logicals vector of the same length as the vector you want to subset. Wherever a True occures, you extra element.

```{r}

lvec <- c(TRUE, FALSE, TRUE) 
x[lvec]

```

You can overwrite elements through subsetting

```{r}

name_vec

name_vec["cate"] <- 10

name_vec

name_vec[1] <- 0

name_vec

name_vec[c(TRUE, FALSE, FALSE, TRUE)] <- 5

name_vec
```

YOu can use negative indices to subset a vector. The negative elements are the ones that are NOT extracted.

```{r}

name_vec[-3]
name_vec[c(-1, -3)]


```

Use [[]] to extract a single element (ex without the name)

```{r}

name_vec[1]
name_vec[[1]]
name_vec[["cat"]]

```

Extract individual names with names function

```{r}

names(name_vec[1])

```

## Exercise

```{r}

x <- c(Yoshi = 10L,
       Mario = 31L,
       Luigi = 72L,
       Peach = 11L,
       Toad  = 38L)

x[c(1, 4)]

x[c(-2:-3,-5)]
#OR
x[-c(2:3,5)]

x[c(TRUE, FALSE, FALSE, TRUE, FALSE)]

x[c("Yoshi", "Peach")]

x [ x %in% c(10,11)]## when want to use corresponding number to call out the elements


```

## Exercise

```{r}

x["Yoshi"] <- 19L
x

x[1] <- 19L

#or

x <- x[sample(1:5)]
x

```

# Recycling vectors

```{r}

x <- c(1, 1, 4, 5, 12)
x +10



```

Interally, R is "recycling" 10 to be the same length as x

```{r}

x + c(10, 10, 10,10, 10)

```

You can "recycle" vectors of lengths greater than 1

```{r}

x

x + c(10,5)


```

Internally, R is 

```{r}

x + c(10, 5, 10, 5, 10)

```

# List

Use 'list" function to create a list

```{r}

my_first_list <- list(x = "a", y = 1, z = c(1L, 2L, 3L), list("a", 1))

my_first_list

```

Usually you use the 'str' function to peak at a vector "structure"

```{r}

str(my_first_list) # more information compare to the first list shown above

```


YOu can use single bracket to create a sublist, but it will still be a list, not the individual elements

```{r}

my_first_list[1] # with name x with value a

my_first_list[c(1,3)]

my_first_list[c(-1,-3)]

my_first_list["y"]

```

Use [[]] to extrac individual elements that are not lists

```{r}

my_first_list[[1]] ## it will get value a instead of "x" with "a"

```

```{r}

data.frame( x = c(1,2,3), y = c(1,2,3)) -> dftemp

dftemp
dftemp[, 1] # do this without tible will change the data frame type

as_tibble(dftemp) -> tibtemp
tibtemp[, 1]

```

Use $ to extract individual elements (just like in a data frame). Will not be a list

```{r}

my_first_list$y
my_first_list$x

```

You can change values through assignment just like in vectorsL

```{r}

my_first_list$x

my_first_list$x <- "b"

str(my_first_list)

```

You can remove elements from a list by setting their values to NULL

```{r}

my_first_list$x <- NULL
str(my_first_list) # then x will be removed, not shown

```

## Exercise

```{r}

wedding <- list(venue = "chick-fil-a",
                guest = tribble(~name,     ~meal, ~age,
                                ##--------/------/-----
                                "Yoshi",   "V",   29L,
                                "Wario",   "C",   27L,
                                "Bowser",  "V",   34L,
                                "Luigi",   "C",   36L,
                                "Toad",    "B",   34L), 
                bride = "Peach",
                groom = "Mario",
                date  = parse_date("11/10/2020", "%d/%m/%Y"))



```

a.
```{r}

wedding$guest%>%
  filter(name != "Wario") -> wedding$guest

wedding$guest

```

b.
```{r}

wedding$meal <- c(v = "Veg",
                  c = "Chicken",
                  B = "Beef")

wedding$meal

```

c.

```{r}

data.frame(wedding)[1]

wedding$venue

wedding[1]

--------------------------

wedding$venue
wedding$date

names(wedding)
wedding[c(1,5)]
wedding[c("venue", "date")]
wedding[c(TRUE, FALSE, FALSE, FALSE, FALSE, TRUE)]


```

d.

```{r}


str_to_sentence(wedding$venue)
str_replace(wedding$venue, "^c", "C") -> wedding$venue

```
































